Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-09 p2/ordering.tex
New partial order that defines ordering

% Define this with the help or a relation / partial order? But then all sequences
% could be ordered while this definition rules out illegal ones.

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple set or sequence of commands $S$,
we use $\orderset{S}$ to denote the set of 
permutations of the commands in $S$
for which the following holds:
for any subsequence $\cxynv\cc\czwmv$,
\begin{itemize}
\item if $n=\parentf{m}$, then $y=\vald$,
\item if $\parentf{n}=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

As usual, by subsequence we mean a subset of the commands in $S$ in the same order as they are found in $S$.
The simplicity of the definition means that sorting the commands
and enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

The principal ideas behind these conditions are 
\cref{ax_separate_commute}, from which we know that commands on incomparable
nodes can be applied in any order, so we only need to focus on the order of commands
with comparable nodes;
\cref{connected_changes}, from which we know that if a simple sequence contains commands on two comparable
nodes, then it contains commands on all nodes in between, so it is enough to specify
the order of commands on directly related nodes;
and finally, \cref{simple_distant_pairs}, from which we know
that commands on parent--child pairs must be construction or destruction pairs.
Investigating these pairs we see that they are identified by the output value of the first command ($y$),
which in turn determines whether the command on the parent or the child must precede the other.

@@

\begin{myclm}\mlabel{subseq_in_orderset}
For simple sequences $S$ and $T$,
if $T\in\orderset{S}$, then for any subsequence of $S$, $S_0$,
and for the corresponding subsequence of $T$, $T_0$, which
contains the same commands but in the order they are found in $T$,
$T_0\in\orderset{S_0}$ must hold.
\end{myclm}
\begin{proof}
This is because
if there are no command pairs in $T$ not conforming to \cref{def_orderset},
then there cannot be any in $T_0$.
\end{proof}

@@

\begin{mycor}\mlabel{order_is_only_possible}
For any simple sequence $S$ and its permutation $S'$, 
\[ S'\not\in\orderset{S} \Longrightarrow S'\equiv\cbrk. \]
\end{mycor}
\begin{proof}
This follows directly from \cref{simple_distant_pairs}.
We use an inverse proof and assume $\wrksx{S'}$.
If $S'\not\in\orderset{S}$ then by \cref{def_orderset}
it contains two commands on a node $n$ and its parent
that, as a subsequence,
do not form a construction or destruction pair.
However, this contradicts \cref{simple_distant_pairs}
which means that no such subsequence exists.
\end{proof}

@@

As a command cannot be both a construction and a destruction command,
this means that the indices $I^S$ are either monotone increasing, or monotone
decreasing, but cannot change direction.
And as $S$ begins with $\cxynv$ and so $I^S_0=1$, we know they are monotone increasing,
and the commands on $n_\indx$ are construction commands.
If we denote the command on $n_\indx$ with $\caaa{x_\indx}{y_\indx}{n_\indx}$,
by definition this means that $[y_\indx]\neq[\empt]$ for all $\indx$.

Let us turn to the location of these commands in $T$,
and denote their indices in $T$ with $I^T_\indx$.
We know $I^T_k=I^T_0-1$ as $\czwmv$ precedes $\cxynv$.
This means there must exist a $\indx$ for which $I^T_{\indx+1}<I^T_\indx$.
We therefore know that
$\caaa{x_{\indx+1}}{y_{\indx+1}}{n_{\indx+1}} \cc \caaa{x_\indx}{y_\indx}{n_\indx}$
is a subsequence of $T$, and by assumption $\parentf{n_{\indx+1}}=n_\indx$.
By definition of $\ordersetsign$, therefore $y_{\indx+1}=\empt$ must be true,
which is a contradiction.

It can be shown in the same way
that the case when the $n_\indx$ nodes ascend the tree,
that is, $n_{\indx+1}=\parentf{n_\indx}$ is true, also leads to a contradiction.
In this case we find that all commands are destruction commands,
and so $[y_\indx]\neq[\vald]$, while in a subsequence of $T$ for a given $\indx$,
$[y_\indx]=[\vald]$ must be true by definition of $\ordersetsign$.
\end{proof}

It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem in some order without breaking it,
this order will be in $\orderset{S}$,
and all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
\begin{mydef}[$\ordp{S}$]
For a simple set or sequence of commands $S$,
$\ordp{S}$ is the unique partial endofunction on filesystems
defined by any order in $\orderset{S}$.
We leave $\ordp{S}$ undefined if such an order does not exist.
\end{mydef}

@@

\begin{mylem}\mlabel{can_simplify}
Each sequence of commands $\wkrsx{S}$
can be transformed into a simple sequence
$T\equiv S$ by applying repeatedly the following transformations:
{\rm(i)} swap neighboring commands on incomparable nodes;
{\rm(ii)} simplify neighboring commands on the same node;
{\rm(iii)} delete assertion commands.
\end{mylem}
\begin{proof}
We can remove all assertion commands, as based on
\cref{ax_assert} the resulting sequence extends the original one.
We therefore assume that $S$ contains no assertion commands.

@@

We therefore know that $S_1\indep\ucx$ and so
from \cref{ax_separate_commute}
\[ S \equiv S_0\cc\ucy\cc\ucx\cc S_1 \cc S_2. \]
We can apply \cref{ax_same_emptyseq,ax_same_singlec} to
$\ucy\cc\ucx$ and get a $S'\eqext S$ that contains one or two less commands on $n$.
Repeating the swap and the simplification we
can convert $S$ into an equivalent simple sequence $S^*$.
\end{proof}

>>>>

We define $\ordersetsign$ with the help of a partial order.
\begin{mydef}[$\orderrel$]\mlabel{def_orderrel}
The binary relation $\orderrel$ holds for some command pairs
on directly related nodes if they are
of the following form:
\begin{itemize}
\item $\caaa{x}{\vald}{\parentf{n}}\orderrel\caaa{z}{w}{n}$, and
\item $\caaa{x}{\empt}{n}\orderrel\caaa{z}{w}{\parentf{n}}$.
\end{itemize}
\end{mydef}
This relation is clearly irreflexive and asymmetric,
and thus defines a partial order.

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple sequence of commands $\wrksx{S}$,
$\orderset{S}$ is the set of permutations satisfying the partial order defined by $\orderrel$.
% For a simple set or sequence of commands $S$,
% we use $\orderset{S}$ to denote the set of 
% permutations of the commands in $S$
% for which the following holds:
% for any subsequence $\cxynv\cc\czwmv$,
% \begin{itemize}
% \item if $n=\parentf{m}$, then $y=\vald$,
% \item if $\parentf{n}=m$, then $y=\empt$.
% \end{itemize}
\end{mydef}

% As usual, by subsequence we mean a subset of the commands in $S$ in the same order as they are found in $S$.
The simplicity of the definitions mean that sorting the commands
and enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

The principal ideas behind the ordering are 
\cref{ax_separate_commute}, from which we know that commands on incomparable
nodes can be applied in any order, so we only need to focus on the order of commands
with comparable nodes;
\cref{connected_changes}, from which we know that if a simple sequence contains commands on two comparable
nodes, then it contains commands on all nodes in between, so it is enough to specify
the order of commands on directly related nodes;
and finally, \cref{simple_distant_pairs}, from which we know
that commands on parent--child pairs must be construction or destruction pairs.
Investigating these pairs we see that they are identified by the output value of the first command ($y$),
which in turn determines whether the command on the parent or the child must precede the other.

@@

\begin{mycor}\mlabel{subseq_in_orderset}
For simple sequences $S$ and $T$,
if $T\in\orderset{S}$, then for any subsequence of $S$, $S_0$,
and for the corresponding subsequence of $T$, $T_0$, which
contains the same commands but in the order they are found in $T$,
$T_0\in\orderset{S_0}$ must hold.
\end{mycor}

@@

\begin{mycor}\mlabel{order_is_only_possible}
For any simple sequence $\wrksx{S}$ and its permutation $S'$, 
\[ S'\not\in\orderset{S} \Longrightarrow S'\equiv\cbrk. \]
\end{mycor}
\begin{proof}
This follows directly from \cref{simple_distant_pairs}.
We use an inverse proof and assume $\wrksx{S'}$.
If $S'\not\in\orderset{S}$ then by \cref{def_orderrel,def_orderset}
it contains two commands on a node $n$ and its parent
that, as a subsequence,
do not form a construction or destruction pair.
However, this contradicts \cref{simple_distant_pairs}
which means that no such subsequence exists.
\end{proof}

@@

As a command cannot be both a construction and a destruction command,
this means that the indices $I^S$ are either monotone increasing, or monotone
decreasing, but cannot change direction.
And as $S$ begins with $\cxynv$ and so $I^S_0=1$, we know they are monotone increasing,
and the commands on $n_\indx$ are construction commands.
If we denote the command on $n_\indx$ with $\caaa{x_\indx}{y_\indx}{n_\indx}$,
by definition this means that
% $[y_\indx]\neq[\empt]$ for all $\indx$.
$y_\indx=\vald$ for all $\indx<k$.

Let us turn to the location of these commands in $T$,
and denote their indices in $T$ with $I^T_\indx$.
We know $I^T_k=I^T_0-1$ as $\czwmv$ precedes $\cxynv$.
This means there must exist a $1\leq \indx<k$ for which $I^T_{\indx+1}<I^T_\indx$.
We therefore know that
$\caaa{x_{\indx+1}}{y_{\indx+1}}{n_{\indx+1}} \cc \caaa{x_\indx}{y_\indx}{n_\indx}$
is a subsequence of $T$.
Because we know $y_\indx=\vald$ and by assumption $\parentf{n_{\indx+1}}=n_\indx$,
this has the form
$\caaa{x_{\indx+1}}{y_{\indx+1}}{n_{\indx+1}} \cc \caaa{x_\indx}{\vald}{\parentf{n_{\indx+1}}}$.
If so, the inverse of $\orderrel$ holds for this subsequence, which is a contradiction.
% By definition of $\ordersetsign$, therefore $y_{\indx+1}=\empt$ must be true,
% which is a contradiction.

It can be shown in the same way
that the case when the $n_\indx$ nodes ascend the tree,
that is, $n_{\indx+1}=\parentf{n_\indx}$ is true, also leads to a contradiction.
In this case we find that all commands are destruction commands,
and so 
% $[y_\indx]\neq[\vald]$, 
$y_\indx=\empt$ for all $1<\indx\leq k$,
which again leads to a subsequence in $T$ violating $\orderrel$.
% while in a subsequence of $T$ for a given $\indx$,
% $[y_\indx]=[\vald]$ must be true by definition of $\ordersetsign$.
\end{proof}

It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem in some order without breaking it,
this order will be in $\orderset{S}$,
and all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
\begin{mydef}[$\ordp{S}$]
For a simple set or sequence of commands $\wrksx{S}$,
$\ordp{S}$ is the unique partial endofunction on filesystems
defined by any order in $\orderset{S}$.
% We leave $\ordp{S}$ undefined if such an order does not exist.
\end{mydef}

@@

\begin{mylem}\mlabel{can_simplify}
Each sequence of commands $\wrksx{S}$
can be transformed into a simple sequence
$T\equiv S$ by applying repeatedly the following transformations:
{\rm(i)} swap neighboring commands on incomparable nodes;
{\rm(ii)} simplify neighboring commands on the same node;
{\rm(iii)} delete assertion commands.
\end{mylem}
\begin{proof}
We can remove all assertion commands, as based on
\cref{ax_assert} the resulting sequence extends the original one.
We therefore assume that $S$ contains no assertion commands.

@@

We therefore know that $S_1\indep\ucx$ and so
from \cref{ax_separate_commute}
\[ S \equiv S_0\cc\ucy\cc\ucx\cc S_1 \cc S_2. \]
We can apply \cref{ax_same_emptyseq,ax_same_singlec} to
$\ucy\cc\ucx$ and get a $S'\eqext S$ that contains one or two less commands on $n$.
Repeating the swap and the simplification we
can convert $S$ into an equivalent simple sequence.
\end{proof}

### 2018-07-09 p2/rec-intro.tex
New partial order that defines ordering

We assume that $A$ and $B$ are returned by an update detector,
and so they are simple sequences.
% TODO
(Otherwise, they can be simplified to become simple sequences; see \cref{update_works}.)
We also assume that $\FS$, $\FS_A$ and $\FS_B$ are not broken.
With these assumptions,
we can now define our reconciliation algorithm.

>>>>

We assume that $A$ and $B$ are returned by an update detector,
and so they are simple sequences.
(Otherwise, they can be simplified to become simple sequences; see \cref{can_simplify}.)
We also assume that $\FS$, $\FS_A$ and $\FS_B$ are not broken.
With these assumptions,
we can now define our reconciliation algorithm.

### 2018-07-09 p2/upd-correct.tex
New partial order that defines ordering

With $\ordersetsign$ we can now define our update detection algorithm.
Its inputs are the original filesystem $\FS$ which has been modified to yield $\FS_1$.
\begin{mydef}[Update detection]\mlabel{def_upddetect}
\begin{enumerate}
    \item For each node $n$ where the value in $\FS$ and $\FS_1$ differ, add the command $\cxynv$ to
        a set of commands $U$.
    \item % TODO
\end{enumerate}
\end{mydef}
% TODO from journal? Simplify sequence?

TODO ENSURE UPDATE DETECTOR IS DEFINED.

@@

\begin{myth}\mlabel{update_works}
For a sequence of commands $U$ returned by the update detector
when comparing the non-broken $\FS^*$ to the original $\FS$,
$U\aFS = \FS^*$.
\end{myth}

TODO What is not trivial here is that $U$ works at all.

TODO And that it can be ordered?

\begin{proof}
Let $\FS$ be the original filesystem, and $\FS^*$
be the filesystem after the changes we intend to detect.
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set of commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
($S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.)
Based on \cref{can_simplify}, there is a simple sequence $S^*$
that is equivalent to $S$.

Therefore we know $S^*\aFS=\FS^*$, and that
$\FS^*$ and $\FS$ differ at exactly those nodes that $S^*$ has commands on.
If $U$ is the simple sequence returned by the update detector
when comparing $\FS^*$ to $\FS$, it must also contain commands on the same set of nodes,
and it must contain the same commands,
as for the command $\cxynv$, $[x]$ must be $[\FS(n)]$, and $y$ must be $\FS^*(n)$.
From this we know that $U$ is a permutation of $S^*$ and so $\orderset{U}=\orderset{S^*}$.

As the update detector uses $\ordersetsign$ to order the commands in $U$, trivially $U\in\orderset{U}$, % TODO
and as $S^*$ is defined on $\FS$,
from \cref{order_is_only_possible} we know $S^* \in \orderset{S^*}$.
This means $U\in\orderset{S^*}$,
and so from \cref{simple_reorder_equiv} we know that $U\equiv S^*$,
from which $U\aFS=\FS^*$.
\end{proof}

>>>>

With $\ordersetsign$ we can now define our update detection algorithm.
Its inputs are the original filesystem $\FS$ which has been modified to yield $\FS^*$.
\begin{mydef}[Update detection]\mlabel{def_upddetect}~
\begin{enumerate}
    \item For each node $n$ where the value in $\FS$ and $\FS^*$ differ, add the command $\caaa{\FS(n)}{\FS^*(n)}{n}$ to
        a set of commands $U_S$. The result is a simple set of commands.
    \item Order the commands according to $\orderrel$, that is, 
        return any sequence $U$ from $\orderset{U_S}$.
\end{enumerate}
\end{mydef}

\Cref{update_works} proves that $U$ functions as expected, that is, $U\aFS=\FS^*$.
While this is trivial if we know that $U$ does not break $\FS$, we
still need to show that it is defined.

On systems where filesystem updates a recorded,
an alternative update detection algorithm is to
omit step 1, and use \cref{can_simplify} to simplify the series of recorded updates into a simple sequence containing the necessary
updates.
In fact, we use such a hypothetical record in the proof \cref{update_works}.

@@

\begin{myth}\mlabel{update_works}
For a simple sequence of commands $U$ returned by the update detector
when comparing the non-broken $\FS^*$ to the original $\FS$,
$U\aFS = \FS^*$.
\end{myth}

\begin{proof}
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set of commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
($S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.)
Based on \cref{can_simplify}, there is a simple sequence $S^*$
that is equivalent to $S$.

Therefore we know $S^*\aFS=\FS^*$, and that
$\FS^*$ and $\FS$ differ at exactly those nodes that $S^*$ has commands on.
$U$ must also contain commands on the same set of nodes,
and it must contain the same commands,
as for the command $\cxynv$, $[x]$ must be $[\FS(n)]$, and $y$ must be $\FS^*(n)$.
From this we know that $U$ is a permutation of $S^*$ and so $\orderset{U}=\orderset{S^*}$.

As the update detector uses $\orderrel$ to order the commands in $U$, trivially $U\in\orderset{U}$,
which means $U\in\orderset{S^*}$,
and so from \cref{simple_reorder_equiv} we know that $U\equiv S^*$,
from which $U\aFS=\FS^*$.
\end{proof}
